Dimension Splitting and a Long Time-Step Multi-Dimensional Scheme for Atmospheric Transport

### 2017-01-24 advectionPaper.tex
Version submitted to QJRMS and to arxiv

%% LyX 2.0.8.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,a4paper,english,british]{qjrms4}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[utf8]{luainputenc}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{color}
\usepackage{array}
\usepackage{verbatim}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{esint}
\usepackage[authoryear]{natbib}

@@

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

@@

@@

\usepackage{babel}
\begin{document}

\title{Dimension Splitting and a Long Time-Step Multi-Dimensional Scheme
for Atmospheric Transport}

\author{Yumeng Chen, Hilary Weller, Stephen Pring and James Shaw}

%\maketitle
\begin{comment}
\tableofcontents{}

\listoftodos
\end{comment}

\begin{abstract}
Dimensionally split advection schemes are attractive for atmospheric
modelling due to their efficiency and accuracy in each spatial dimension.
Accurate long time steps can be achieved without significant cost
using the flux-form semi-Lagrangian technique. The dimensionally split
scheme used in this paper is constructed from the one-dimensional
Piecewise Parabolic Method and extended to two dimensions using COSMIC
splitting. The dimensionally split scheme is compared with a genuinely
multi-dimensional, method of lines scheme with implicit time-stepping
which is stable for very large Courant numbers.

@@

Such mesh distortions are expected to accentuate the errors associated
with dimension splitting, however, the dimensionally split scheme
is very accurate on orthogonal meshes and accuracy decreases only
a little in the presence of large mesh distortions. The dimensionally
split scheme retains accuracy when long time-steps are used. The multi-dimensional
scheme is almost entirely insensitive to mesh distortions and asymptotes
to second-order accuracy at high resolution. As is expected for implicit
time-stepping, phase errors occur when using long time-steps but the
spatially well resolved features are advected at the correct speed
and the multi-dimensional scheme is always stable.

A naive estimate of computational cost (number of multiplies) reveals
that the implicit scheme is  the most expensive, particularly for
large Courant numbers. If the multi-dimensional scheme is used instead
with explicit time-stepping, the Courant number is restricted to less
than one but the cost becomes similar to the dimensionally split scheme.
\end{abstract}
\begin{keywords}
hello
\end{keywords}

@@

\begin{figure}
\noindent \begin{centering}
\scalebox{0.5}[0.5]{\input{figs/cubicUpCPCstencil.pstex_t}}\hfill{}\scalebox{0.5}[0.5]{\input{figs/hexStencil.pstex_t}}
\par\end{centering}

@@

\begin{figure}
\begin{centering}
\includegraphics[width=0.49\textwidth,height=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh}
\par\end{centering}

@@

\begin{figure}
\begin{tabular}{cc}
Orthogonal, dimensionally split scheme & Non-orthogonal, dimensionally split scheme\tabularnewline
\includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_1_Tcontours} & \includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_1_Tcontours}\tabularnewline
Orthogonal, multi-dimensional scheme & Non-orthogonal, multi-dimensional scheme\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_1_Tcontours} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_1_Tcontours}\tabularnewline
\multicolumn{2}{c}{\includegraphics[width=0.9\linewidth]{HilarysGraphics/solidBodyRotationOnPlane_legends_Tdiff}}\tabularnewline
\end{tabular}

@@

\begin{figure}
\begin{tabular}{cc}
Orthogonal, dimensionally split scheme & Non-orthogonal, dimensionally split scheme\tabularnewline
\includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_10_TcontoursBig} & \includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_10_TcontoursBig}\tabularnewline
Orthogonal, multi-dimensional scheme & Non-orthogonal, multi-dimensional scheme\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_10_TcontoursBig} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_10_TcontoursBig}\tabularnewline
\multicolumn{2}{c}{\includegraphics[width=0.9\linewidth]{HilarysGraphics/solidBodyRotationOnPlane_legends_TdiffBig_Tdiff}}\tabularnewline
\end{tabular}

\caption{Results of the solid body rotation test case on orthogonal and non-orthogonal
meshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=10\text{ s}$
giving a maximum Courant number of around 10. The contours show the
tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
The dashed contour shows the initial conditions. The min and max text
in each plot gives the minimum and maximum errors. \label{fig:solidBody-dt10}}
\end{figure}

@@

\begin{figure}
\begin{centering}
\begin{tabular}{cc}
$\ell_{2}$ errors, convergence with $\Delta x$ for $c\approx1$ & $\ell_{\infty}$ errors, convergence with $\Delta x$ for $c\approx1$\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_l2errorsC1} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_linferrorsC1}\tabularnewline
$\ell_{2}$ errors, convergence with $\Delta x$ for $c\approx10$ & $\ell_{\infty}$ errors, convergence with $\Delta x$ for $c\approx10$\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_l2errorsC10} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_linferrorsC10}\tabularnewline
\end{tabular}
\par\end{centering}

@@

\begin{figure}
\begin{centering}
\begin{tabular}{cc}
$\ell_{2}$ errors & $\ell_{\infty}$ errors\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_l2errors_dt} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_linferrors_dt}\tabularnewline
\end{tabular}
\par\end{centering}

\caption{Convergence with temporal resolution of the $\ell_{2}$ and $\ell_{\infty}$
errors for the solid body rotation test case at $t=500\ \text{seconds}$
($5/6$ of one revolution) using orthogonal and non-orthogonal meshes
of $100\times100$ cells using multi-dimensional and dimensionally
split schemes. Maximum Courant numbers range from $c=0.25$ (for $\Delta t=0.25\ \text{s})$
to $ $$c=10$ (for $\Delta t=10\ \text{s})$. \label{fig:solidBodyConverge-dt} }
\end{figure}

@@

All simulations use basic terrain following co-ordinates \citep[BTF,][]{GCS75}
in order to present a challenging test case that maximises the non-orthogonality.
The transformation is given by:
\begin{equation}
\begin{array}{ccc}
X=x &  & Z=H\frac{z-h(x)}{H-h(x)}\end{array}
\end{equation}
where $H$ is the domain height and $h$ is the terrain height. The
test case uses a domain of width 300\ km, height, $H=25\text{ km}$
and a mountain range defined by
\begin{equation}
h=\begin{cases}
h_{0}\cos^{2}\frac{\pi x}{\lambda}\cos^{2}\frac{\pi x}{2a} & \text{ for }|x|\le a\\
0 & \text{ otherwise}
\end{cases}
\end{equation}
with the maximum mountain height, $h_{0}=3\text{ km}$, half-width
$a=25\text{ km}$ and wavelength $\lambda=8\text{ km}$. These values
give a maximum terrain gradient of close to $45^{o}$. The wind is
given by a streamfunction which is defined at vertices so that the
wind field is discretely divergence free. The streamfunction at vertices
is calculated analytically from the wind profile:
\begin{equation}
u(z)=u_{0}\begin{cases}
1 & \text{ for }z_{2}\le z\\
\sin^{2}\left(\frac{\pi}{2}\frac{z-z_{1}}{z_{2}-z_{1}}\right) & \text{ for }z_{1}<z\le z_{2}\\
0 & \text{ for }z<z_{1}
\end{cases}
\end{equation}
with $u_{0}=10\text{ ms}^{-1}$, $z_{1}=4\text{ km}$ and $z_{2}=5\text{ km}$.
The initial tracer position is given by:
\begin{equation}
\phi=\begin{cases}
\cos^{2}\frac{\pi r}{2} & \text{ for }r\le1\\
0 & \text{ otherwise}
\end{cases}\text{ with }r=\sqrt{\left(\frac{x-x_{0}}{A_{x}}\right)^{2}+\left(\frac{z-z_{0}}{A_{z}}\right)^{2}}
\end{equation}
with initial tracer centre, $(x_{0},z_{0})=(-50\text{ km},\ 9\text{ km})$
and halfwidths $A_{x}=25\text{ km}$, $A_{z}=9\text{ km}$. At time
$t=5000\text{ s}$ the tracer is above the mountain and the simulation
finishes at $t=10,000\text{ s}$ by which time the analytic solution
is centred at $(50\text{km },8\text{km})$.
\begin{figure}
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.14\textwidth}>{\centering}p{0.42\textwidth}c}
 & Dimensionally split & Multi-dimensional\tabularnewline
$\Delta t=25\ $s

max $c$ = 0.74

horiz $c$ =0.25

$c_{d}=0.22$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt25CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt25CN_up_10000_Tall}\tabularnewline
$\Delta t=50\ $s

max $c$ = 1.48

horiz $c$ =0.5

$c_{d}=0.44$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt50CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt50CN_up_10000_Tall}\tabularnewline
$\Delta t=100\ $s

max $c$ = 2.96

horiz $c$ =1

$c_{d}=0.87$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt100CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt100CN_up_10000_Tall}\tabularnewline
$\Delta t=200\ $s

max $c$ = 5.93

horiz $c$ =2

$c_{d}=1.75$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt200CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt200CN_up_10000_Tall}\tabularnewline
$\Delta t=500\ $s

max $c$ = 14.8

horiz $c$ =5

$c_{d}=4.4$ & unstable & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt500CN_up_10000_Tall}\tabularnewline
$\Delta t=1000\ $s

max $c$ = 29.6

horiz $c$ =10

$c_{d}=8.7$ & unstable & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt1000CN_up_10000_Tall}\tabularnewline
\multicolumn{3}{c}{\includegraphics[width=0.85\textwidth]{HilarysGraphics/vSlice_horizontalAdvection_orography_legends_Tall_TdiffAll}}\tabularnewline
\end{tabular*}

@@

\begin{figure}
\noindent \begin{centering}
\begin{tabular}{cc}
$\ell_{2}$ error & $\ell_{\infty}$ error\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/orography_l2errors} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/orography_linferrors}\tabularnewline
\end{tabular}
\par\end{centering}

\caption{Convergence of the $\ell_{2}$ and $\ell_{\infty}$ error metrics
for horizontal advection over a 3000m and a 6000m mountains after
10,000s for the dimensionally split (PPM) scheme and multi-dimensional
MOL scheme, all using a horizontal Courant number of 0.25. Lines also
show theoretical first- and second-order convergence are also shown.
\label{fig:horizontalAdvectConverge}}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\textwidth]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_120x60_0_TUmesh}

\includegraphics[width=1\textwidth]{HilarysGraphics/deformingAdvectionCyclic_legends_Tmesh_T}

@@

\begin{figure}
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.03\textwidth}cc}
$t$ & Dimensionally split & Multi-dimensional\tabularnewline
1 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T1} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_1_T}\tabularnewline
2 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T2} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_2_T}\tabularnewline
3 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T3} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_3_T}\tabularnewline
4 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T4} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_4_T}\tabularnewline
5 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T5} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_5_T}\tabularnewline
\multicolumn{3}{c}{\includegraphics[width=0.98\textwidth]{HilarysGraphics/deformingAdvectionCyclic_legends_T}}\tabularnewline
\end{tabular*}

@@

\begin{figure}
\begin{tabular}{cc}
$\ell_{2}$ error, $c\le1$ & $\ell_{\infty}$ error, $c\le1$\tabularnewline
\includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_l2errors_c1} & \includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_linferrors_c1}\tabularnewline
$\ell_{2}$ error, $c\le10$ & $\ell_{\infty}$ error, $c\le10$\tabularnewline
\includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_l2errors_c10} & \includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_linferrors_c10}\tabularnewline
\end{tabular}

@@

\begin{figure}
\begin{tabular}{cc}
$\ell_{2}$ error & $\ell_{\infty}$ error\tabularnewline
\includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_l2errors_dt} & \includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_linferrors_dt}\tabularnewline
\end{tabular}

>>>>

%% LyX 2.0.8.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,a4paper,english,british,times,doublespace]{qjrms4}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{esint}
\usepackage[authoryear]{natbib}

@@

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

@@

\usepackage{lineno}
%\linenumbers

@@

\usepackage{babel}
\begin{document}
\title{Dimension Splitting and a Long Time-Step Multi-Dimensional Scheme for Atmospheric Transport}
\author{Yumeng Chen,\affil{a} Hilary Weller,\affil{b}\corrauth\ Stephen Pring\affil{c} and James Shaw\affil{b}}

\runningheads{Y. Chen, H. Weller, S. Pring and J. Shaw}{Dimension splitting for Advection}

\address{
\affilnum{a}Universit\"at Hamburg\\
\affilnum{b}Meteorology, University of Reading \\
\affilnum{c}Met Office, UK}
\corraddr{E-mail: <h.weller@reading.ac.uk>}

\begin{abstract}\normalfont
Dimensionally split advection schemes are attractive for atmospheric modelling due to their efficiency and accuracy in each spatial dimension. Accurate long time steps can be achieved without significant cost using the flux-form semi-Lagrangian technique. The dimensionally split scheme used in this paper is constructed from the one-dimensional Piecewise Parabolic Method and extended to two dimensions using COSMIC splitting. The dimensionally split scheme is compared with a genuinely multi-dimensional, method of lines scheme with implicit time-stepping which is stable for very large Courant numbers.

@@

Such mesh distortions are expected to accentuate the errors associated with dimension splitting, however, the dimensionally split scheme is very accurate on orthogonal meshes and accuracy decreases only a little in the presence of large mesh distortions. The dimensionally split scheme also loses some accuracy when long time-steps are used. The multi-dimensional scheme is almost entirely insensitive to mesh distortions and asymptotes to second-order accuracy at high resolution. As is expected for implicit time-stepping, phase errors occur when using long time-steps but the spatially well resolved features are advected at the correct speed and the multi-dimensional scheme is always stable.

A naive estimate of computational cost (number of multiplies) reveals that the implicit scheme is  the most expensive, particularly for large Courant numbers. If the multi-dimensional scheme is used instead with explicit time-stepping, the Courant number is restricted to less than one but the cost becomes similar to the dimensionally split scheme.
\end{abstract}

\keywords{
Multi-dimensional; advection; stability; accuracy; long time-steps
}
\maketitle

%\tableofcontents
%\listoftodos
\onecolumn

@@

\begin{figure}
\noindent \begin{centering}
\includegraphics{stencils}
\par\end{centering}

@@

\begin{figure}
\begin{centering}
\includegraphics[width=0.49\textwidth,height=0.49\textwidth]{solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh}
\par\end{centering}

@@

\begin{figure}
\includegraphics{sbrc1.png}

@@

\begin{figure}
\includegraphics{sbrc10.png}

\caption{Results of the solid body rotation test case on orthogonal and non-orthogonalmeshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=10\text{ s}$ giving a maximum Courant number of around 10. The contours show the tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors. The dashed contour shows the initial conditions. The min and max text in each plot gives the minimum and maximum errors. \label{fig:solidBody-dt10}}
\end{figure}

@@

\begin{figure}
\begin{centering}
\includegraphics{sbr_dx}
\par\end{centering}

@@

\begin{figure}
\begin{centering}
\includegraphics{sbr_dt}
\par\end{centering}

\caption{Convergence with temporal resolution of the $\ell_{2}$ and $\ell_{\infty}$errors for the solid body rotation test case at $t=500\ \text{seconds}$ ($5/6$ of one revolution) using orthogonal and non-orthogonal meshes of $100\times100$ cells using multi-dimensional and dimensionally split schemes. Maximum Courant numbers range from $c=0.25$ (for $\Delta t=0.25\ \text{s})$ to $ $$c=10$ (for $\Delta t=10\ \text{s})$. \label{fig:solidBodyConverge-dt} }
\end{figure}

@@

All simulations use basic terrain following co-ordinates \citep[BTF,][]{GCS75}
in order to present a challenging test case that maximises the non-orthogonality.
The transformation is given by:
\begin{equation}
\begin{array}{ccc}
X=x &  & Z=H\frac{z-h(x)}{H-h(x)}\end{array}
\end{equation}
where $H$ is the domain height and $h$ is the terrain height. The
test case uses a domain of width 300\ km, height, $H=25\text{ km}$
and a mountain range defined by
\begin{equation}
h=\begin{cases}
h_{0}\cos^{2}\frac{\pi x}{\lambda}\cos^{2}\frac{\pi x}{2a} & \text{ for }|x|\le a\\
0 & \text{ otherwise}
\end{cases}
\end{equation}
with the maximum mountain height, $h_{0}=3\text{ km}$, half-width
$a=25\text{ km}$ and wavelength $\lambda=8\text{ km}$. These values
give a maximum terrain gradient of close to $45^{o}$. The wind is
given by a streamfunction which is defined at vertices so that the
wind field is discretely divergence free. The streamfunction at vertices
is calculated analytically from the wind profile:
\begin{equation}
u(z)=u_{0}\begin{cases}
1 & \text{ for }z_{2}\le z\\
\sin^{2}\left(\frac{\pi}{2}\frac{z-z_{1}}{z_{2}-z_{1}}\right) & \text{ for }z_{1}<z\le z_{2}\\
0 & \text{ for }z<z_{1}
\end{cases}
\end{equation}
with $u_{0}=10\text{ ms}^{-1}$, $z_{1}=4\text{ km}$ and $z_{2}=5\text{ km}$.
The initial tracer position is given by:
\begin{equation}
\phi=\begin{cases}
\cos^{2}\frac{\pi r}{2} & \text{ for }r\le1\\
0 & \text{ otherwise}
\end{cases}\text{ with }r=\sqrt{\left(\frac{x-x_{0}}{A_{x}}\right)^{2}+\left(\frac{z-z_{0}}{A_{z}}\right)^{2}}
\end{equation}
with initial tracer centre, $(x_{0},z_{0})=(-50\text{ km},\ 9\text{ km})$
and halfwidths $A_{x}=25\text{ km}$, $A_{z}=9\text{ km}$. At time
$t=5000\text{ s}$ the tracer is above the mountain and the simulation
finishes at $t=10,000\text{ s}$ by which time the analytic solution
is centred at $(50\text{km },8\text{km})$.

\begin{figure}
\includegraphics{overMountains.png}

@@

\begin{figure}
\noindent \begin{centering}
\includegraphics{overMountains_dx}
\par\end{centering}

\caption{Convergence of the $\ell_{2}$ and $\ell_{\infty}$ error metricsfor horizontal advection over a 3000m and a 6000m mountains after 10,000s for the dimensionally split (PPM) scheme and multi-dimensional MOL scheme, all using a horizontal Courant number of 0.25. Lines also show theoretical first- and second-order convergence are also shown.
\label{fig:horizontalAdvectConverge}}
\end{figure}

@@

\begin{figure}
\includegraphics{deform_init}

@@

\begin{figure}
\includegraphics{deform.png}

@@

\begin{figure}
\includegraphics{deform_dx}

@@

\begin{figure}
\includegraphics{deform_dt}
